IML Term project

### 2022-12-05 main_markdown.rmd
Refactoring, brought in tree methods.
Wrote some more stuff to paper.rmd.

Columns selected by correlation

@@

# Divide training data into two for training & testing
set.seed(42)
idx <- sample(nrow(npf), nrow(npf)/2)
npf_train <- npf[idx, ]
npf_test <- npf[-idx, ]

@@

```{r functions}
# mean squared error
mse <- function(yhat, y) sqrt(mean((y - yhat)**2))

@@

# Find cross-validation predictions
cv <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               n = nrow(data), # number of rows in the data matrix
               k = min(n, 10), # number of cross-validation folds
               split = kpart(n, k), # the split of n data items into k folds
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:k) {
        ## go through all folds, train on other folds, and make a prediction
        mod <- train(data[split != i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[split == i] <- pred(mod, data[split == i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
# loocv <- function(
#                formula, # Formula specifying which variables to use
#                data, # Dataset
#                model = lm, # Type of model to train (as a function)
#                ## function to train a model on data
#                train = function(data) model(formula, data = data),
#                ## function to make predictions on the trained model
#                pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
#     yhat <- NULL
#     for (i in 1:nrow(data)) {
#         ## go through all stations, train on other stations, and make a prediction
#         mod <- train(data[-i, ])
#         if (is.null(yhat)) {
#             ## initialise yhat to something of correct data type,
#             yhat <- pred(mod, data)
#         } else {
#             yhat[i] <- pred(mod, data[i, ])
#         }
#     }
#     yhat # finally, output cross-validation predictions
# }
```

@@

```{r naive bayes}
library(e1071)
nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))
# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))
```

### LDA
```{r lda}
library(MASS)
lda.fit <- lda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)
# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))
lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))
```

### QDA
```{r qda}
library(MASS)
qda.fit <- qda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)
# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))
# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))
```

### GLM
```{r glm, warning=FALSE}
mod1<-glm(class2~., npf, family=binomial(link="logit"))
mod2<-glm(class2~., npf, family=binomial(link="probit"))
plot(mod1)
prob1<-predict(mod1, npf, type="response")
pred1<-c()
perp1<-c()
for(i in 1:dim(npf)[1]){
  if(prob1[i]>=0.5){
    pred1<-c(pred1,1)
  } else{
    pred1<-c(pred1,0)
  }
  if(npf$class2[i]==1){
    perp1<-c(perp1,prob1[i])
  } else{
    perp1<-c(perp1,1-prob1[i])
  }
}
mean(pred1==npf$class2)
exp(-mean(perp1))
```

### Lasso / Ridge
```{r lassoridge, message=FALSE}
library(glmnet)
mod3<-cv.glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=1, family="binomial", type.measure="class")
mod4<-cv.glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=0, family="binomial", type.measure = "class")
mod3a<-glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=1, family="binomial", lambda=mod3$lambda.min)
coef(mod3)
plot(mod3a)
plot(mod3$glmnet.fit, xvar="lambda")
prob3a<-predict(mod3a, newx=as.matrix(npf[,1:100]), type="response")
pred3a<-c()
perp3a<-c()
for(i in 1:dim(npf)[1]){
  if(prob3a[i]>=0.5){
    pred3a<-c(pred3a,1)
  } else{
    pred3a<-c(pred3a,0)
  }
  if(npf$class2[i]==1){
    perp3a<-c(perp3a,prob3a[i])
  } else{
    perp3a<-c(perp3a,1-prob3a[i])
  }
}
mean(pred3a==npf$class2)
exp(-mean(perp3a))
mod4a<-glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=0, family="binomial", lambda=mod4$lambda.min)
plot(mod4a)
plot(mod4$glmnet.fit, xvar="lambda")
prob4a<-predict(mod4a, newx=as.matrix(npf[,1:100]), type="response")
pred4a<-c()
perp4a<-c()
for(i in 1:dim(npf)[1]){
  if(prob4a[i]>=0.5){
    pred4a<-c(pred4a,1)
  } else{
    pred4a<-c(pred4a,0)
  }
  if(npf$class2[i]==1){
    perp4a<-c(perp4a,prob4a[i])
  } else{
    perp4a<-c(perp4a,1-prob4a[i])
  }
}
mean(pred4a==npf$class2)
exp(-mean(perp4a))
```

### SVM
```{r svm, warning=FALSE}
modsv<-tune(svm, class2~., data=npf.fact, kernel="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1, 2, 5)))
modsvm<-svm(class2~., data=npf.fact, kernel="linear", scale=FALSE, cost=0.001,probability=TRUE)
predsv<-predict(modsvm,newdata=npf.fact)
predsv2<-predict(modsvm,newdata=npf.fact,probability = T)
probsv<-attr(predsv2, "probabilities")
perpsv<-c()
for(i in 1:dim(npf)[1]){
  if(npf$class2[i]==0){
    perpsv<-c(perpsv, probsv[i,2])
  } else{
    perpsv<-c(perpsv, probsv[i,1])
  }
  }
mean(predsv==npf$class2)
exp(-mean(log(perpsv)))
```

@@

@@

```{r correlation at different heights, message=FALSE}
# Korrelaatiomatriisi
cm <- cor(npf[, endsWith(colnames(npf), ".mean")])
# Remove .mean from column and row names.
colnames(cm) <- rownames(cm) <- sapply(colnames(cm), function(s) gsub(".mean", "", s))
# Muuttujakohtaiset korrelaatiot (kun enemmän kuin yksi muuttuja, jolla sama alkuosa)
kable(cm[startsWith(colnames(cm), "CO2"), startsWith(colnames(cm), "CO2")],
      caption = "Correlation (CO2)") %>%
  kable_styling()
kable(cm[startsWith(colnames(cm), "H2O"), startsWith(colnames(cm), "H2O")],
      caption = "Correlation (H20)") %>%
  kable_styling()
kable(cm[startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NO)") %>%
  kable_styling()
kable(cm[startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NOx)") %>%
  kable_styling()
kable(cm[startsWith(colnames(cm), "O3"), startsWith(colnames(cm), "O3")],
      caption = "Correlation (O3)") %>%
  kable_styling()
kable(cm[startsWith(colnames(cm), "RHIRGA"), startsWith(colnames(cm), "RHIRGA")],
      caption = "Correlation (RHIRGA)") %>%
  kable_styling()
kable(cm[startsWith(colnames(cm), "T"), startsWith(colnames(cm), "T")],
      caption = "Correlation (T)") %>%
  kable_styling()
```

@@

The measurement height 16.8m is the only one with all measurements. Because of this and the correlation between the measurements at different heights, we choose to discard measurements heights other than 16.8m.

```{r muuttujakorkeudet, message=FALSE}
# Tunnistetaan, miltä korkeudelta on otettu mitäkin mittauksia
# Sarakkeet
cols <- unname(sapply(colnames(npf[endsWith(colnames(npf), ".mean")]), function(s) gsub(".mean", "", s)))
# Taulukko muuttujista eri korkeuksissa (vain niille muuttujille, joista mittauksia eri korkeuksissa)
kable(data.frame(dm.42 = c(sort(cols[endsWith(cols, "42")])[1:6], NA, sort(cols[endsWith(cols, "42")])[7]),
                                dm.84 = c(NA, sort(cols[endsWith(cols, "84")])[1:5], NA, sort(cols[endsWith(cols, "84")])[6]),
                                dm.168 = c(sort(cols[endsWith(cols, "168")])),
                                dm.336 = c(sort(cols[endsWith(cols, "336")])[1:4], NA, sort(cols[endsWith(cols, "336")])[5], NA, NA),
                                dm.504 = c(sort(cols[endsWith(cols, "504")])[1:6], NA, sort(cols[endsWith(cols, "504")])[7]),
                                dm.672 = c(NA, NA, sort(cols[endsWith(cols, "672")])[2:5], NA, sort(cols[endsWith(cols, "672")])[6])),
      caption = "Measurements at different heights: What measurements have been done at particular heights?") %>%
  kable_styling()
```

>>>>

@@

# Divide numerical training data into two for training & testing
set.seed(42)
idx <- sample(nrow(npf), nrow(npf)/2)
npf_train <- npf[idx, ]
npf_test <- npf[-idx, ]

@@

```{r functions}
## root mean squared error
rmse <- function(yhat, y) sqrt(mean((y - yhat)**2))

## Classification accuracy
cacc <- function(pred, true) sum(as.integer(as.logical(pred == true)))/length(pred)

@@

# Find cross-validation predictions
cv <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               n = nrow(data), # number of rows in the data matrix
               k = min(n, 10), # number of cross-validation folds
               split = kpart(n, k), # the split of n data items into k folds
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:k) {
        ## go through all folds, train on other folds, and make a prediction
        mod <- train(data[split != i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[split == i] <- pred(mod, data[split == i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
loocv <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:nrow(data)) {
        ## go through all stations, train on other stations, and make a prediction
        mod <- train(data[-i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[i] <- pred(mod, data[i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
```

@@

```{r naive bayes}
library(e1071)
nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran
nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))
# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))
kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### LDA
```{r lda}
library(MASS)
lda.fit <- lda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)
# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))
lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))
kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))%>%
  kable_styling(latex_options = "HOLD_position")
```

### QDA
```{r qda}
library(MASS)
qda.fit <- qda(class2 ~ . , data = npf_train)
# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)
# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}
qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))
# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))
kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### GLM
```{r glm, warning=FALSE, fig.height=4}
mod1<-glm(class2~., npf, family=binomial(link="logit"))
mod2<-glm(class2~., npf, family=binomial(link="probit"))
plot(mod1)
prob1<-predict(mod1, npf, type="response")
pred1<-c()
perp1<-c()
for(i in 1:dim(npf)[1]){
  if(prob1[i]>=0.5){
    pred1<-c(pred1,1)
  } else{
    pred1<-c(pred1,0)
  }
  if(npf$class2[i]==1){
    perp1<-c(perp1,prob1[i])
  } else{
    perp1<-c(perp1,1-prob1[i])
  }
}
```

GLM accuracy and perplexity. 
```{r glm res, warning=FALSE}
mean(pred1==npf$class2)
exp(-mean(perp1))
```

### Lasso / Ridge
```{r lasso, message=FALSE, fig.height=4}
library(glmnet)
mod3<-cv.glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=1, family="binomial", type.measure="class")
mod4<-cv.glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=0, family="binomial", type.measure = "class")
mod3a<-glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=1, family="binomial", lambda=mod3$lambda.min)
coef(mod3)
plot(mod3a)
plot(mod3$glmnet.fit, xvar="lambda")
prob3a<-predict(mod3a, newx=as.matrix(npf[,1:100]), type="response")
pred3a<-c()
perp3a<-c()
for(i in 1:dim(npf)[1]){
  if(prob3a[i]>=0.5){
    pred3a<-c(pred3a,1)
  } else{
    pred3a<-c(pred3a,0)
  }
  if(npf$class2[i]==1){
    perp3a<-c(perp3a,prob3a[i])
  } else{
    perp3a<-c(perp3a,1-prob3a[i])
  }
}
```

Lasso accuracy and perplexity. 
```{r lasso  res, warning=FALSE}
mean(pred3a==npf$class2)
exp(-mean(perp3a))
```

```{r ridge, message=FALSE, fig.height=4}
mod4a<-glmnet(x=as.matrix(npf[,1:100]), y=npf[,101],alpha=0, family="binomial", lambda=mod4$lambda.min)
plot(mod4a)
plot(mod4$glmnet.fit, xvar="lambda")
prob4a<-predict(mod4a, newx=as.matrix(npf[,1:100]), type="response")
pred4a<-c()
perp4a<-c()
for(i in 1:dim(npf)[1]){
  if(prob4a[i]>=0.5){
    pred4a<-c(pred4a,1)
  } else{
    pred4a<-c(pred4a,0)
  }
  if(npf$class2[i]==1){
    perp4a<-c(perp4a,prob4a[i])
  } else{
    perp4a<-c(perp4a,1-prob4a[i])
  }
}
```

Ridge accuracy and perplexity. 
```{r ridge res, warning=FALSE}
mean(pred4a==npf$class2)
exp(-mean(perp4a))
```

### SVM
```{r svm, warning=FALSE}
modsv<-tune(svm, class2~., data=npf.fact, kernel="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1, 2, 5)))
modsvm<-svm(class2~., data=npf.fact, kernel="linear", scale=FALSE, cost=0.001,probability=TRUE)
predsv<-predict(modsvm,newdata=npf.fact)
predsv2<-predict(modsvm,newdata=npf.fact,probability = T)
probsv<-attr(predsv2, "probabilities")
perpsv<-c()
for(i in 1:dim(npf)[1]){
  if(npf$class2[i]==0){
    perpsv<-c(perpsv, probsv[i,2])
  } else{
    perpsv<-c(perpsv, probsv[i,1])
  }
}
```
SVM accuracy and perplexity. 
```{r SVM res, warning=FALSE}
mean(predsv==npf$class2)
exp(-mean(log(perpsv)))
```

### Tree-based methods

```{r trees, message=FALSE, warning=FALSE}
library(tree)
library(MASS)
library(randomForest) # random forest
library(gbm) # GBM
library (ISLR2)

#' Cross-validation predictor (from week 1 problem 2)¨
cv <- function(
    formula, # variables to use
    data, # data
    model = lm, # model to train
    n = nrow(data), # rows in the matrix
    k = min(n, 10), # cross-validation folds
    split = kpart(n, k), # the split of n data items into k folds
    train = function(data) model(formula, data = data),
    pred = function(model, data) predict(model, newdata = data)) {
  yhat <- NULL
  for (i in 1:k) {
    mod <- train(data[split != i, ])
    if (is.null(yhat)) {
      yhat <- pred(mod, data)
    } else {
      yhat[split == i] <- pred(mod, data[split == i, ])
    }
  }
  return(yhat)
}

loocv <- function(
    formula, # Formula specifying which variables to use
    data, # Dataset
    model = lm, # Type of model to train (as a function)
    ## function to train a model on data
    train = function(data) model(formula, data = data),
    ## function to make predictions on the trained model
    pred = function(model, data) predict(model, newdata = data)) {
  yhat <- NULL
  for (i in 1:nrow(data)) {
    ## go through all stations, train on other stations, and make a prediction
    mod <- train(data[-i, ])
    if (is.null(yhat)) {
      ## initialise yhat to something of correct data type,
      yhat <- pred(mod, data)
    } else {
      yhat[i] <- pred(mod, data[i, ])
    }
  }
  yhat # finally, output cross-validation predictions
}

#' Dummy model (from week 1 problem 2)
dummy <- function(formula, data) {
  target <- all.vars(formula[[2]])
  lm(as.formula(sprintf("%s ~ 1", target)), data)
}

#' Models
models <- list(
  Dummy = dummy,
  RandomForest = randomForest,
  GBM = gbm
)

# gmb_pred <- predict(gbm(class2 ~ ., data = npf, cv.folds = 10, train.fraction = 0.5), data=npf)
# print(cacc(gmb_pred, npf$class2))

#' Errors
errors <- sapply(models, function(model) {
  mod <- model(class2 ~ ., data = npf_train)
  c(
    c(
      train = rmse(predict(mod, data = npf_train), npf_train$class2),
      test = rmse(predict(mod, data = npf_test), npf_test$class2),
      CV = rmse(cv(mod, data = npf_train), npf_train$class2),
      LOOCV = rmse(loocv(mod, data = npf_train), npf_train$class2) 
    )
  )
})

#' Classification accuracies
results <- sapply(models, function(model) {
  mod <- model(class2 ~ ., data = npf_train)
  c(
    c(
      train = cacc(predict(mod, data = npf_train), npf_train$class2),
      test = cacc(predict(mod, data = npf_test), npf_test$class2),
      CV = cacc(cv(mod, data = npf_train), npf_train$class2),
      LOOCV = cacc(loocv(mod, data = npf_train), npf_train$class2)
    )
  )
})

kable(errors,
      caption = "Tree-based methods, errors") %>%
  kable_styling(latex_options = "HOLD_position")

kable(results,
      caption = "Tree-based methods, classification accuracy") %>%
  kable_styling(latex_options = "HOLD_position")
```

\newpage

@@

\newpage

@@

```{r correlation at different heights, message=FALSE}
# Korrelaatiomatriisi
cm <- cor(npf[, endsWith(colnames(npf), ".mean")])
# Remove .mean from column and row names.
colnames(cm) <- rownames(cm) <- sapply(colnames(cm), function(s) gsub(".mean", "", s))
# Muuttujakohtaiset korrelaatiot (kun enemmän kuin yksi muuttuja, jolla sama alkuosa)
kable(cm[startsWith(colnames(cm), "CO2"), startsWith(colnames(cm), "CO2")],
      caption = "Correlation (CO2)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "H2O"), startsWith(colnames(cm), "H2O")],
      caption = "Correlation (H20)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NO)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NOx)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "O3"), startsWith(colnames(cm), "O3")],
      caption = "Correlation (O3)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "RHIRGA"), startsWith(colnames(cm), "RHIRGA")],
      caption = "Correlation (RHIRGA)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "T"), startsWith(colnames(cm), "T")],
      caption = "Correlation (T)") %>%
  kable_styling(latex_options = "HOLD_position")
```

\newpage

@@

The measurement height 16.8m is the only one with all measurements. Because of this and the correlation between the measurements at different heights, we choose to discard measurements from heights other than 16.8m.

```{r muuttujakorkeudet, message=FALSE}
# Tunnistetaan, miltä korkeudelta on otettu mitäkin mittauksia
# Sarakkeet
cols <- unname(sapply(colnames(npf[endsWith(colnames(npf), ".mean")]), function(s) gsub(".mean", "", s)))
# Taulukko muuttujista eri korkeuksissa (vain niille muuttujille, joista mittauksia eri korkeuksissa)
kable(data.frame(dm.42 = c(sort(cols[endsWith(cols, "42")])[1:6], NA, sort(cols[endsWith(cols, "42")])[7]),
                                dm.84 = c(NA, sort(cols[endsWith(cols, "84")])[1:5], NA, sort(cols[endsWith(cols, "84")])[6]),
                                dm.168 = c(sort(cols[endsWith(cols, "168")])),
                                dm.336 = c(sort(cols[endsWith(cols, "336")])[1:4], NA, sort(cols[endsWith(cols, "336")])[5], NA, NA),
                                dm.504 = c(sort(cols[endsWith(cols, "504")])[1:6], NA, sort(cols[endsWith(cols, "504")])[7]),
                                dm.672 = c(NA, NA, sort(cols[endsWith(cols, "672")])[2:5], NA, sort(cols[endsWith(cols, "672")])[6])),
      caption = "Measurements at different heights: What measurements have been done at particular heights?") %>%
  kable_styling(latex_options = "HOLD_position")
```
